### 2022-03-29 Draft Report/Draft Project.Rmd
Created little template

---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Insert title of project here"
subtitle: "Web address for GitHub repository"
author: "Name"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory

# Load your packages
library(agricolae)
library(corrplot)
library(cowplot)
library(dataRetrieval)
library(extrafont)
library(extrafontdb)
library(ggpubr)
library(ggthemes)
library(hrbrthemes)
library(Kendall)
library(leaflet)
library(lubridate)
library(mapview)
library(rvest)
library(sf)
library(tidycensus)
library(tidyverse)
library(trend)
library(tseries)
library(viridis)
library(zoo)

# Set your ggplot theme
#Disable on-the-fly projections
sf::sf_use_s2(FALSE)
#Fix Mapview
mapviewOptions(fgb = FALSE)

Theme2 <- theme_ipsum()+
  theme(legend.position = "bottom",
        legend.key = element_rect(fill = "white", colour = "black"),legend.direction = "horizontal",
        legend.title = element_text(face = "bold"))
theme_set(Theme2)

# Load your datasets

```

# Rationale and Research Questions

\newpage

# Dataset Information

\newpage

# Exploratory Analysis

\newpage

# Analysis

## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2:

\newpage

# Summary and Conclusions

\newpage

# References
<add references here if relevant, otherwise delete this section>

### 2022-03-29 Draft Report/Draft Project.Rmd
added working directory, change as you wish

@@
 
+```{r, initialize}
+# Set your working directory
+getwd()
+knitr::opts_knit$set(root.dir = "~/Desktop/Duke MPP/Environ Data /Environmental_Data_Analytics_2022/")
+
+```
+
 ```{r setup, include=FALSE}
-# Set your working directory
-
 # Load your packages

### 2022-03-30 Draft Report/Draft Project.Rmd
Update Draft Project.Rmd

@@
 mainfont: Times New Roman
-
 ---
 
+```{=tex}
 \newpage
@@
 \newpage
-
+```
 ```{r, initialize}
@@
 
+## Potential research questions!!!
+
+-   Write what you're thinking about!
+
 \newpage
@@
 
-## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>
+## Question 1: \<insert specific question here and add additional subsections for additional questions below, if needed>
 
@@
 # References
-<add references here if relevant, otherwise delete this section>
+
+\<add references here if relevant, otherwise delete this section>

### 2022-04-05 Draft Report/Draft Project.Rmd
draft
two links
Updating draft

@@
 ---
+
+## Potential research questions!!!
+
+-   Write what you're thinking about!
+-   <https://www.thetransportpolitic.com/2022/01/24/openings-and-construction-starts-planned-for-2022/>
+- <https://theintercept.com/2022/02/12/wildfires-prisons-climate-california/>
+- <https://cseveren.github.io/files/racialized_difference_commuting_20220320.pdf>
+<https://github.com/OpenFloodAI/Climate-Change-Datasets>
+https://github.com/firstlookmedia/climate-and-punishment-data
+\newpage-
 
@@
 
-## Potential research questions!!!
+# Dataset Information
+## Methodology
+### Gather Data Sources
+####EJSCREEN NC
+Gathered From: 
+#### National Risk Index
+Gathered from: https://hazards.fema.gov/nri/data-resources North Carolina
 
--   Write what you're thinking about!
+### Clean Data Sources
+Filter and pipe each to just the variables of interest: location, titles, variables 1 and 2
 
-\newpage
+### Present Descriptive Statistics and Maps
 
-# Dataset Information
-
+###
 \newpage

### 2022-04-05 Draft Report/Draft Project.Rmd
added a few notes

--- 2022-03-30 Draft Report/Draft Project.Rmd
@@
 ---
+
+## Potential research questions!!!
+
+-   Write what you're thinking about!
+-   <https://www.thetransportpolitic.com/2022/01/24/openings-and-construction-starts-planned-for-2022/>
+-   <https://theintercept.com/2022/02/12/wildfires-prisons-climate-california/>
+-   <https://cseveren.github.io/files/racialized_difference_commuting_20220320.pdf> <https://github.com/OpenFloodAI/Climate-Change-Datasets> <https://github.com/firstlookmedia/climate-and-punishment-data> \newpage-
 
@@
 
-## Potential research questions!!!
+# Dataset Information
 
--   Write what you're thinking about!
+## Methodology
 
-\newpage
+### Gather Data Sources
 
-# Dataset Information
+####EJSCREEN Gathered From: 
+2020
+#### National Risk Index 
+Gathered from: <https://hazards.fema.gov/nri/data-resources> North Carolina
+
+### Clean Data Sources
+
+Filter and pipe each to just the variables of interest: location, titles, variables 1 and 2
+
+Questions:(NC, FL, CA)
+1. 1960 Housing (size of datapoint) v Coastal Flood risk (Color on map)
+2. Percent Low Income v  Flood Risk
+--3? PM2.5 (size of datapoint) v Wildfire Risk  (color)--
+### Present Descriptive Statistics and Maps
+### Time Series
 

### 2022-04-10 Draft Report/Draft Project.Rmd
Made progress on filtering

@@
 ---
+```{r, initialize}
+# Set your working directory
+getwd()
+knitr::opts_knit$set(root.dir = '/xxxDesktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject')
 
+```
 ## Potential research questions!!!
@@
 \newpage
-```
-```{r, initialize}
-# Set your working directory
-getwd()
-knitr::opts_knit$set(root.dir = "~/Desktop/Duke MPP/Environ Data /Environmental_Data_Analytics_2022/")
-
 ```
@@
 ### Gather Data Sources
+```{r load datasources}
+NRICali <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/NRI_Table_CensusTracts_California/NRI_Table_CensusTracts_California.csv', stringsAsFactors = TRUE)
+NRINorCar <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/NRI_Table_CensusTracts_NorthCarolina/NRI_Table_CensusTracts_NorthCarolina.csv', stringsAsFactors = TRUE)
+NRIFlor <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/NRI_Table_CensusTracts_Florida/NRI_Table_CensusTracts_Florida.csv', stringsAsFactors = TRUE)
+EJScreen21 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/EJSCREEN_2021_StatePctile_Tracts.csv')
+```
 
@@
 Filter and pipe each to just the variables of interest: location, titles, variables 1 and 2
+```{r initial filter}
+NRICali2 <- NRICali %>%
+  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,
+         ends_with("RATNG"),ends_with("RISKS"),ends_with("RISKR"),ends_with("EVENTS"),
+         ends_with("SCORE"),ends_with("EALS"),ends_with("EALR")) %>%
+         select(!starts_with("LTNG"),!starts_with("AVLN"),!starts_with("HAIL"),
+                !starts_with("LNDS"),!starts_with("VLCN"))
+
+NRIFlor2 <- NRIFlor %>%
+  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,
+         ends_with("RATNG"),ends_with("RISKS"),ends_with("RISKR"),ends_with("EVENTS"),
+         ends_with("SCORE"),ends_with("EALS"),ends_with("EALR")) %>%
+         select(!starts_with("LTNG"),!starts_with("AVLN"),!starts_with("HAIL"),
+                !starts_with("LNDS"),!starts_with("VLCN"))
+
+NRINorCar2 <- NRINorCar %>%
+  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,
+         ends_with("RATNG"),ends_with("RISKS"),ends_with("RISKR"),ends_with("EVENTS"),
+         ends_with("SCORE"),ends_with("EALS"),ends_with("EALR")) %>%
+         select(!starts_with("LTNG"),!starts_with("AVLN"),!starts_with("HAIL"),
+                !starts_with("LNDS"),!starts_with("VLCN"))
+
+EJScreen21 <- EJScreen21 %>%
+  filter(grepl('Florida|California|North Carolina',STATE_NAME))
+```
+
+```{r risk filters}
+#flood risks
+NRICaliFlood <- NRICali2 %>%
+  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))
+NRIFlorFlood <- NRIFlor2 %>%
+  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))
+NRINorCarFlood <- NRINorCar2 %>%
+  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))
+
+#fire risks
+NRICaliFire <- NRICali2 %>%
+  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("WFIR"))
+NRIFlorFire <- NRIFlor2 %>%
+  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("WFIR"))
+NRINorCarFire <- NRINorCar2 %>%
+  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("WFIR"))
+```
 

### 2022-04-10 Draft Report/Draft Project.Rmd
Updated the filter process

@@
 theme_set(Theme2)
-
-# Load your datasets
-
 ```
@@
 ```{r load datasources}
+# Load your datasets
 NRICali <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/NRI_Table_CensusTracts_California/NRI_Table_CensusTracts_California.csv', stringsAsFactors = TRUE)
@@
 NRIFlor <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/NRI_Table_CensusTracts_Florida/NRI_Table_CensusTracts_Florida.csv', stringsAsFactors = TRUE)
-EJScreen21 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/EJSCREEN_2021_StatePctile_Tracts.csv')
+EJScreen21 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/EJSCREEN_2021_StatePctile_Tracts.csv',stringsAsFactors = TRUE)
 ```
@@
 
+Questions:(NC, FL, CA)
+1. 1960 Housing (size of datapoint) v Coastal Flood risk (Color on map)
+2. Percent Low Income v  Flood Risk
+--3? PM2.5 (size of datapoint) v Wildfire Risk  (color)--
 ```{r risk filters}
@@
 
-Questions:(NC, FL, CA)
-1. 1960 Housing (size of datapoint) v Coastal Flood risk (Color on map)
-2. Percent Low Income v  Flood Risk
---3? PM2.5 (size of datapoint) v Wildfire Risk  (color)--
+```{r EJScreen Filters}
+EJ_filter <- EJScreen21 %>%
+  select(ID,STATE_NAME,contains("INCPCT"),contains("HSPCT"),LINGISOPCT,PRE1960PCT,contains("LDPNT"),contains("PM25"))
+```
+
 ### Present Descriptive Statistics and Maps
+
 ### Time Series

### 2022-04-10 Draft Report/Draft Project.Rmd
changes

@@
 geometry: margin=2.54cm
-title: "Insert title of project here"
+title: "Risk of Coastal Flooding"
 subtitle: "Web address for GitHub repository"
@@
 
+```{r}
+
+mydata <- readr::read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/NRI_Table_CensusTracts_NorthCarolina/NRI_Table_CensusTracts_NorthCarolina.csv')
+
+mymap <-st_read('/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/NRI_Table_CensusTracts_NorthCarolina/NRI_Table_CensusTracts_NorthCarolina.csv')
+
+str(mymap)
+
 ### Time Series

### 2022-04-11 Draft Report/Draft Project.Rmd
lets see

--- 2022-04-10 Draft Report/Draft Project.Rmd
@@
 getwd()
-knitr::opts_knit$set(root.dir = '/xxxDesktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject')
+knitr::opts_knit$set(root.dir = '/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject')
 
@@
 # Load your datasets
-NRICali <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/NRI_Table_CensusTracts_California/NRI_Table_CensusTracts_California.csv', stringsAsFactors = TRUE)
-NRINorCar <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/NRI_Table_CensusTracts_NorthCarolina/NRI_Table_CensusTracts_NorthCarolina.csv', stringsAsFactors = TRUE)
-NRIFlor <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/NRI_Table_CensusTracts_Florida/NRI_Table_CensusTracts_Florida.csv', stringsAsFactors = TRUE)
-EJScreen21 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Data Sources/EJSCREEN_2021_StatePctile_Tracts.csv',stringsAsFactors = TRUE)
+NRICali <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_California/NRI_Table_CensusTracts_California.csv', stringsAsFactors = TRUE)
+NRINorCar <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_NorthCarolina/NRI_Table_CensusTracts_NorthCarolina.csv', stringsAsFactors = TRUE)
+NRIFlor <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_Florida/NRI_Table_CensusTracts_Florida.csv', stringsAsFactors = TRUE)
+EJScreen21 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/EJSCREEN_2021_StatePctile_Tracts.csv',stringsAsFactors = TRUE)
 ```
@@
   select(ID,STATE_NAME,contains("INCPCT"),contains("HSPCT"),LINGISOPCT,PRE1960PCT,contains("LDPNT"),contains("PM25"))
+
 ```
@@
 ### Present Descriptive Statistics and Maps
+Add in Shape Files 
+https://www.census.gov/geographies/mapping-files/time-series/geo/cartographic-boundary.html
+```{r census tracts}
+california_tract <- st_read('./Potenital Data Sources/cb_2020_cali_tract_500k/cb_2020_06_tract_500k.shp')
+norcar_tract <- st_read('./Potential Data Sources/cb_2020_norcar_tract_500k/cb_2020_37_tract_500k.shp')
+florida_tract <- st_read('./Potential Data Sources/cb_2020_flor_tract_500k/cb_2020_12_tract_500k.shp')
+```
 

### 2022-04-11 Draft Report/Draft Project.Rmd
Added EJ Screen Data Download source

@@
 NRIFlor <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_Florida/NRI_Table_CensusTracts_Florida.csv', stringsAsFactors = TRUE)
-EJScreen21 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/EJSCREEN_2021_StatePctile_Tracts.csv',stringsAsFactors = TRUE)
+EJScreen20 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/EJSCREEN_2020_StatePctile.csv',stringsAsFactors = TRUE)
+colnames(EJScreen20)
+#colnames(EJScreen21)
+#EJScreen21 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/EJSCREEN_2021_StatePctile_Tracts.csv',stringsAsFactors = TRUE)
 ```
@@
 Gathered from: <https://hazards.fema.gov/nri/data-resources> North Carolina
-
+And from <https://gaftp.epa.gov/EJSCREEN/2020/> for the EJ Screen Data
 ### Clean Data Sources
@@
 
-EJScreen21 <- EJScreen21 %>%
+EJScreen20 <- EJScreen20 %>%
   filter(grepl('Florida|California|North Carolina',STATE_NAME))
@@
 ```{r EJScreen Filters}
-EJ_filter <- EJScreen21 %>%
+EJ_filter <- EJScreen20 %>%
   select(ID,STATE_NAME,contains("INCPCT"),contains("HSPCT"),LINGISOPCT,PRE1960PCT,contains("LDPNT"),contains("PM25"))

### 2022-04-12 Draft Report/Draft Project.Rmd
At the point where both datasets can be mapped against each other for california, see line 229 for the "good result"

@@
 getwd()
-knitr::opts_knit$set(root.dir = '/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject')
-
+knitr::opts_knit$set(root.dir = '/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Project/')
+getwd()
 ```
@@
 
+EJ_California <- EJ_filter %>%
+    filter(STATE_NAME=="California")
+EJ_NorCar <- EJScreen20 %>%
+  filter(STATE_NAME=="North Carolina")
+
+EJ_Florida <-EJScreen20 %>%
+  filter(STATE_NAME=="Florida")
+#write.csv(EJ_NorCar,'./Potential Data Sources/Census Tracts/EJScreen_NC.csv')
+#write.csv(EJ_Florida,'./Potential Data Sources/Census Tracts/EJScreen_FLorida.csv')
 ```
@@
 ```{r census tracts}
-california_tract <- st_read('./Potenital Data Sources/cb_2020_cali_tract_500k/cb_2020_06_tract_500k.shp')
-norcar_tract <- st_read('./Potential Data Sources/cb_2020_norcar_tract_500k/cb_2020_37_tract_500k.shp')
-florida_tract <- st_read('./Potential Data Sources/cb_2020_flor_tract_500k/cb_2020_12_tract_500k.shp')
+california_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_cali_tract_500k/cb_2020_06_tract_500k.shp',)
+norcar_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_norcar_tract_500k/cb_2020_37_tract_500k.shp')
+florida_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_flor_tract_500k/cb_2020_12_tract_500k.shp')
+#mapview(california_tract,zcol="AWATER")
+```
+
+```{r california file filtering}
+california_tract <- california_tract %>%
+  select(GEOID,NAMELSAD,NAMELSADCO,ALAND,AWATER,geometry) %>%
+  mutate(TRACTFIPS = as.double(substr(california_tract$GEOID,2,15)))
+NRICaliFire <- NRICaliFire %>%
+  select(TRACTFIPS,POPULATION,starts_with("WFIR"))
+NRICaliFlood <- NRICaliFlood %>%
+  select(TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))
+
+# See GEOID Structure, <https://www.census.gov/programs-surveys/geography/guidance/geo-identifiers.html> all the other stuff is Census Tracts not BlockGroups
+names(EJ_California)[names(EJ_California) == 'ID'] <- 'BLOCKID'
+EJ_California <- EJ_California %>%
+  mutate(TRACTID = as.double(substr(EJ_California$BLOCKID,1,10)))
+
+#now we convert into averages for the whole census tract. Imprecise but we've gotta do it
+EJ_Cali <- EJ_California %>%
+  select(!starts_with("T_")) %>%
+  group_by(TRACTID) %>%
+  summarize(across(!where(is.factor),mean))
+
+```
+https://www.census.gov/geographies/reference-files/time-series/geo/gazetteer-files.2020.html
+```{r longitude latitute for counties}
+#https://www.census.gov/geographies/reference-files/time-series/geo/gazetteer-files.2020.html
+gazetteer <- read.delim('./Potential Data Sources/Census Tracts/2021_Gaz_tracts_national.txt',stringsAsFactors = TRUE)
+
+ca_gazetteer <- gazetteer %>%
+  filter(USPS=="CA")%>%
+  select(!c(ALAND_SQMI,AWATER_SQMI))%>%
+  mutate(TRACTID = ca_gazetteer$GEOID)
+```
+
+```{r attempts to merge}
+CA_NRI_fire_shp <- left_join(california_tract,NRICaliFire,by="TRACTFIPS")
+CA_NRI_flood_shp <- left_join(california_tract,NRICaliFlood,by="TRACTFIPS")
+
+CA_EJ_df <- left_join(ca_gazetteer,EJ_Cali,by="TRACTID")
+CA_EJ_shp <- CA_EJ_df %>%
+  st_as_sf(coords = c('INTPTLONG','INTPTLAT'), crs=4269)
+# mapview(CA_NRI_shp)
+mapview(CA_EJ_shp)
+```
+
+```{r mapview tests}
+mapview(list(CA_NRI_fire_shp,CA_EJ_shp),zcol=c("WFIR_RISKS","PRE1960PCT"))
+
+mapview(list(CA_NRI_flood_shp,CA_EJ_shp),
+        zcol=list("CFLD_RISKS","PRE1960PCT"),
+        layer.name=list("Coastal Flood Risk","Relative Building Ages"),
+        col.regions=list(mapviewPalette("mapviewVectorColors"),mapviewPalette("mapviewTopoColors")),
+        alpha.regions=.5)
+
+mapview(CA_NRI_fire_shp,zcol="WFIR_RISKS")
+mapview(CA_EJ_shp,zcol="LOWINCPCT",alpha.regions=.5,col.regions= mapviewPalette("mapviewTopoColors"))
+mapview(CA_EJ_shp,zcol="PRE1960PCT")
+
+mapshot(m, file = paste0(getwd(), "/map.png"),
+          remove_controls = c("homeButton", "layersControl"))
 ```

### 2022-04-12 Draft Report/Draft Project.Rmd
Updated with a bit more statistical analysis

@@
   st_as_sf(coords = c('INTPTLONG','INTPTLAT'), crs=4269)
-# mapview(CA_NRI_shp)
-mapview(CA_EJ_shp)
+```
+
+```{r simple graphs}
+ggplot(CA_EJ_df,aes(y=LOWINCPCT,x=PM25))+geom_point(aes(color=LINGISOPCT))+geom_smooth(method="lm")+labs(title="PM25 v Low Income %",y="% Low Income Census Tract",x="PM25 Concentration")
+
+ggplot(CA_NRI_fire_shp,aes(y=POPULATION,x=WFIR_RISKR))+geom_boxplot()+coord_flip()+labs(title="Fire Risk versus Population",y="Fire Risk")
+ggplot(CA_NRI_flood_shp,aes(y=POPULATION,x=CFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Coastal Flood Risk versus Population",y="Coastal Flood Risk")
+ggplot(CA_NRI_flood_shp,aes(y=POPULATION,x=RFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Riverine Flood Risk versus Population",y="Riverine Flood Risk")
+```
+
+```{r simple mapview tests}
+mapview(CA_NRI_fire_shp,zcol="POPULATION")+ mapview(CA_NRI_flood_shp,zcol="RFLD_RISKR") +mapview(CA_EJ_shp,zcol="LOWINCPCT")
 ```
@@
 ```{r mapview tests}
-mapview(list(CA_NRI_fire_shp,CA_EJ_shp),zcol=c("WFIR_RISKS","PRE1960PCT"))
+mapview(list(CA_NRI_fire_shp,CA_EJ_shp),
+        zcol=c("WFIR_RISKS","PRE1960PCT"),
+        layer.name=list("Fire Risk","Relative Building Ages"))
 
@@
 
-mapshot(m, file = paste0(getwd(), "/map.png"),
-          remove_controls = c("homeButton", "layersControl"))
+# mapshot(m, file = paste0(getwd(), "/map.png"),
+#           remove_controls = c("homeButton", "layersControl"))
 ```

### 2022-04-18 Draft Report/Draft Project.Rmd
Updated Draft REport to Include NC and FL NRI and Demographics
6 Shape files now exist, NRI Fire and Flood for both FL and NC, as well as the shape files for dot points with demographic info for NC/FL based on EJScreen data

--- 2022-04-11 Draft Report/Draft Project.Rmd
@@
 ---
+
 ```{r, initialize}
@@
 getwd()
-knitr::opts_knit$set(root.dir = '/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject')
+knitr::opts_knit$set(root.dir = '/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Project/')
+getwd()
+```
 
-```
 ## Potential research questions!!!
@@
 ```
-
 ```{r setup, include=FALSE}
@@
 ### Gather Data Sources
+
 ```{r load datasources}
@@
 EJScreen20 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/EJSCREEN_2020_StatePctile.csv',stringsAsFactors = TRUE)
-colnames(EJScreen20)
+#colnames(EJScreen20)
 #colnames(EJScreen21)
@@
 
-####EJSCREEN Gathered From: 
-2020
-#### National Risk Index 
-Gathered from: <https://hazards.fema.gov/nri/data-resources> North Carolina
-And from <https://gaftp.epa.gov/EJSCREEN/2020/> for the EJ Screen Data
-### Clean Data Sources
+####EJSCREEN Gathered From: 2020 #### National Risk Index Gathered from: <https://hazards.fema.gov/nri/data-resources> North Carolina And from <https://gaftp.epa.gov/EJSCREEN/2020/> for the EJ Screen Data ##\# Clean Data Sources
 
 Filter and pipe each to just the variables of interest: location, titles, variables 1 and 2
+
 ```{r initial filter}
@@
 
-Questions:(NC, FL, CA)
-1. 1960 Housing (size of datapoint) v Coastal Flood risk (Color on map)
-2. Percent Low Income v  Flood Risk
---3? PM2.5 (size of datapoint) v Wildfire Risk  (color)--
+Questions:(NC, FL, CA) 1. 1960 Housing (size of datapoint) v Coastal Flood risk (Color on map) 2. Percent Low Income v Flood Risk --3? PM2.5 (size of datapoint) v Wildfire Risk (color)--
+
 ```{r risk filters}
@@
 
+EJ_California <- EJ_filter %>%
+    filter(STATE_NAME=="California")
+
+EJ_NorCar <- EJ_filter %>%
+  filter(STATE_NAME=='North Carolina')
+
+EJ_Florida <-EJ_filter %>%
+  filter(STATE_NAME=="Florida")
+
+# EJ_NorCar <- EJScreen20 %>%
+#   filter(STATE_NAME=='North Carolina')
+# 
+# EJ_Florida <-EJScreen20 %>%
+#   filter(STATE_NAME=="Florida")
+
+#write.csv(EJ_NorCar,'./Potential Data Sources/Census Tracts/EJScreen_NC.csv',col.names)
+#write.csv(EJ_Florida,'./Potential Data Sources/Census Tracts/EJScreen_FLorida.csv')
 ```
@@
 ### Present Descriptive Statistics and Maps
-Add in Shape Files 
-https://www.census.gov/geographies/mapping-files/time-series/geo/cartographic-boundary.html
+
+Add in Shape Files <https://www.census.gov/geographies/mapping-files/time-series/geo/cartographic-boundary.html>
+
 ```{r census tracts}
-california_tract <- st_read('./Potenital Data Sources/cb_2020_cali_tract_500k/cb_2020_06_tract_500k.shp')
-norcar_tract <- st_read('./Potential Data Sources/cb_2020_norcar_tract_500k/cb_2020_37_tract_500k.shp')
-florida_tract <- st_read('./Potential Data Sources/cb_2020_flor_tract_500k/cb_2020_12_tract_500k.shp')
+california_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_cali_tract_500k/cb_2020_06_tract_500k.shp',)
+norcar_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_norcar_tract_500k/cb_2020_37_tract_500k.shp')
+florida_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_flor_tract_500k/cb_2020_12_tract_500k.shp')
+#mapview(california_tract,zcol="AWATER")
+```
+
+```{r california file filtering}
+california_tract <- california_tract %>%
+  select(GEOID,NAMELSAD,NAMELSADCO,ALAND,AWATER,geometry) %>%
+  mutate(TRACTFIPS = as.double(substr(california_tract$GEOID,2,15)))
+NRICaliFire <- NRICaliFire %>%
+  select(TRACTFIPS,POPULATION,starts_with("WFIR"))
+NRICaliFlood <- NRICaliFlood %>%
+  select(TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))
+
+# See GEOID Structure, <https://www.census.gov/programs-surveys/geography/guidance/geo-identifiers.html> all the other stuff is Census Tracts not BlockGroups
+names(EJ_California)[names(EJ_California) == 'ID'] <- 'BLOCKID'
+EJ_California <- EJ_California %>%
+  mutate(TRACTID = as.double(substr(EJ_California$BLOCKID,1,10)))
+
+#now we convert into averages for the whole census tract. Imprecise but we've gotta do it
+EJ_Cali <- EJ_California %>%
+  select(!starts_with("T_")) %>%
+  group_by(TRACTID) %>%
+  summarize(across(!where(is.factor),mean))
+```
+
+```{r florida file filtering}
+florida_tract <- florida_tract %>%
+  select(GEOID,NAMELSAD, NAMELSADCO, ALAND, AWATER, geometry) %>%
+  mutate(TRACTFIPS = as.double(substr(florida_tract$GEOID, 1,15)))
+NRIFlorFire <-NRIFlorFire %>%
+  select(TRACTFIPS,POPULATION,starts_with ("WFIR"))
+NRIFlorFlood <- NRIFlorFlood %>%
+  select(TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))
+
+EJScreen_Florida <- EJ_Florida
+names(EJScreen_Florida)[names(EJScreen_Florida) == 'ID'] <- 'BLOCKID' 
+EJScreen_Florida <- EJScreen_Florida %>% mutate(TRACTID = as.double(substr(EJScreen_Florida$BLOCKID,1,11)))
+
+EJ_Flor <- EJScreen_Florida %>%
+select(!starts_with("T_")) %>%
+group_by(TRACTID) %>%
+summarize(across(!where(is.factor), mean))
+```
+
+```{r north carolina file filtering}
+norcar_tract <- norcar_tract %>%
+  select(GEOID,NAMELSAD, NAMELSADCO, ALAND, AWATER, geometry) %>%
+  mutate(TRACTFIPS = as.double(substr(norcar_tract$GEOID, 1,15)))
+NRINorCarFire <-NRINorCarFire %>%
+  select(TRACTFIPS,POPULATION,starts_with ("WFIR"))
+NRINorCarFlood <- NRINorCarFlood %>%
+  select(TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))
+
+EJScreen_norcar <- EJ_NorCar
+names(EJScreen_norcar)[names(EJScreen_norcar) == 'ID'] <- 'BLOCKID' 
+EJScreen_norcar <- EJScreen_norcar %>% mutate(TRACTID = as.double(substr(EJScreen_norcar$BLOCKID,1,11)))
+
+EJ_NorCar <- EJScreen_norcar %>%
+select(!starts_with("T_")) %>%
+group_by(TRACTID) %>%
+summarize(across(!where(is.factor), mean))
+```
+
+<https://www.census.gov/geographies/reference-files/time-series/geo/gazetteer-files.2020.html>
+
+```{r longitude latitute for counties}
+#https://www.census.gov/geographies/reference-files/time-series/geo/gazetteer-files.2020.html
+gazetteer <- read.delim('./Potential Data Sources/Census Tracts/2021_Gaz_tracts_national.txt',stringsAsFactors = TRUE)
+
+ca_gazetteer <- gazetteer %>%
+  filter(USPS=="CA")
+
+ca_gazetteer <- ca_gazetteer %>%  
+  select(!c(ALAND_SQMI,AWATER_SQMI))%>%
+  mutate(TRACTID = ca_gazetteer$GEOID)
+
+fl_gazetteer <- gazetteer %>%
+  filter(USPS=="FL")
+
+fl_gazetteer <- fl_gazetteer %>%
+  select(!c(ALAND_SQMI, AWATER_SQMI)) %>%
+  mutate(TRACTID = fl_gazetteer$GEOID)
+
+nc_gazetteer <- gazetteer %>%
+  filter(USPS=="NC")
+
+nc_gazetteer <- nc_gazetteer %>%
+  select(!c(ALAND_SQMI, AWATER_SQMI)) %>%
+  mutate(TRACTID = nc_gazetteer$GEOID)
+```
+
+```{r attempts to merge}
+#California
+CA_NRI_fire_shp <- left_join(california_tract,NRICaliFire,by="TRACTFIPS")
+CA_NRI_flood_shp <- left_join(california_tract,NRICaliFlood,by="TRACTFIPS")
+
+CA_EJ_df <- left_join(ca_gazetteer,EJ_Cali,by="TRACTID")
+CA_EJ_shp <- CA_EJ_df %>%
+  st_as_sf(coords = c('INTPTLONG','INTPTLAT'), crs=4269)
+
+#Florida
+FL_NRI_fire_shp <- left_join(florida_tract,NRIFlorFire,by="TRACTFIPS")
+FL_NRI_flood_shp <- left_join(florida_tract,NRIFlorFlood,by="TRACTFIPS")
+
+FL_EJ_df <- left_join(fl_gazetteer,EJ_Flor,by="TRACTID")
+FL_EJ_shp <- FL_EJ_df %>%
+  st_as_sf(coords = c('INTPTLONG','INTPTLAT'), crs=4269)
+
+#North Carolina
+NC_NRI_fire_shp <- left_join(norcar_tract,NRINorCarFire,by="TRACTFIPS")
+NC_NRI_flood_shp <- left_join(norcar_tract,NRINorCarFlood,by="TRACTFIPS")
+
+NC_EJ_df <- left_join(nc_gazetteer,EJ_NorCar,by="TRACTID")
+NC_EJ_shp <- NC_EJ_df %>%
+  st_as_sf(coords = c('INTPTLONG','INTPTLAT'), crs=4269)
+```
+
+```{r simple graphs CA}
+ggplot(CA_EJ_df,aes(y=LOWINCPCT,x=PM25))+geom_point(aes(color=LINGISOPCT))+geom_smooth(method="lm")+labs(title="PM25 v Low Income %",y="% Low Income Census Tract",x="PM25 Concentration")
+
+ggplot(CA_NRI_fire_shp,aes(y=POPULATION,x=WFIR_RISKR))+geom_boxplot()+coord_flip()+labs(title="Fire Risk versus Population",y="Fire Risk")
+ggplot(CA_NRI_flood_shp,aes(y=POPULATION,x=CFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Coastal Flood Risk versus Population",y="Coastal Flood Risk")
+ggplot(CA_NRI_flood_shp,aes(y=POPULATION,x=RFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Riverine Flood Risk versus Population",y="Riverine Flood Risk")
+```
+
+```{r simple graphs FL}
+ggplot(FL_EJ_df,aes(y=LOWINCPCT,x=PM25))+geom_point(aes(color=LINGISOPCT))+geom_smooth(method="lm")+labs(title="PM25 v Low Income %",y="% Low Income Census Tract",x="PM25 Concentration")
+
+ggplot(FL_NRI_fire_shp,aes(y=POPULATION,x=WFIR_RISKR))+geom_boxplot()+coord_flip()+labs(title="Fire Risk versus Population",y="Fire Risk")
+ggplot(FL_NRI_flood_shp,aes(y=POPULATION,x=CFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Coastal Flood Risk versus Population",y="Coastal Flood Risk")
+ggplot(FL_NRI_flood_shp,aes(y=POPULATION,x=RFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Riverine Flood Risk versus Population",y="Riverine Flood Risk")
+```
+
+```{r simple graphs NC}
+ggplot(NC_EJ_df,aes(y=LOWINCPCT,x=PM25))+geom_point(aes(color=LINGISOPCT))+geom_smooth(method="lm")+labs(title="PM25 v Low Income %",y="% Low Income Census Tract",x="PM25 Concentration")
+
+ggplot(NC_NRI_fire_shp,aes(y=POPULATION,x=WFIR_RISKR))+geom_boxplot()+coord_flip()+labs(title="Fire Risk versus Population",y="Fire Risk")
+ggplot(NC_NRI_flood_shp,aes(y=POPULATION,x=CFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Coastal Flood Risk versus Population",y="Coastal Flood Risk")
+ggplot(NC_NRI_flood_shp,aes(y=POPULATION,x=RFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Riverine Flood Risk versus Population",y="Riverine Flood Risk")
+```
+
+```{r simple mapview tests}
+#CA
+mapview(CA_NRI_fire_shp,zcol="POPULATION")+ mapview(CA_NRI_flood_shp,zcol="RFLD_RISKR") +mapview(CA_EJ_shp,zcol="LOWINCPCT")
+#comment
+#FL
+mapview(FL_NRI_fire_shp,zcol="POPULATION")+ mapview(FL_NRI_flood_shp,zcol="RFLD_RISKR") +mapview(FL_EJ_shp,zcol="LOWINCPCT")
+
+#NC
+mapview(NC_NRI_fire_shp,zcol="POPULATION")+mapview(NC_NRI_flood_shp,zcol="RFLD_RISKR")+mapview(NC_EJ_shp,zcol="LOWINCPCT")
+```
+
+```{r mapview tests CA}
+mapview(list(CA_NRI_fire_shp,CA_EJ_shp),
+        zcol=c("WFIR_RISKS","PRE1960PCT"),
+        layer.name=list("Fire Risk","Relative Building Ages"))
+
+mapview(list(CA_NRI_flood_shp,CA_EJ_shp),
+        zcol=list("CFLD_RISKS","PRE1960PCT"),
+        layer.name=list("Coastal Flood Risk","Relative Building Ages"),
+        col.regions=list(mapviewPalette("mapviewVectorColors"),mapviewPalette("mapviewTopoColors")),
+        alpha.regions=.5)
+
+mapview(CA_NRI_fire_shp,zcol="WFIR_RISKS")
+mapview(CA_EJ_shp,zcol="LOWINCPCT",alpha.regions=.5,col.regions= mapviewPalette("mapviewTopoColors"))
+mapview(CA_EJ_shp,zcol="PRE1960PCT")
+
+# mapshot(m, file = paste0(getwd(), "/map.png"),
+#           remove_controls = c("homeButton", "layersControl"))
 ```

### 2022-04-18 Draft Report/Draft Project.Rmd
updated file path structure to be malleable

@@
 # Load your datasets
-NRICali <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_California/NRI_Table_CensusTracts_California.csv', stringsAsFactors = TRUE)
-NRINorCar <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_NorthCarolina/NRI_Table_CensusTracts_NorthCarolina.csv', stringsAsFactors = TRUE)
-NRIFlor <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_Florida/NRI_Table_CensusTracts_Florida.csv', stringsAsFactors = TRUE)
-EJScreen20 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/EJSCREEN_2020_StatePctile.csv',stringsAsFactors = TRUE)
+home_path <- '~/Desktop/Duke MPP/Environ Data /872 Final Project'
+
+NRICali <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_California/NRI_Table_CensusTracts_California.csv',sep=""), stringsAsFactors = TRUE)
+NRINorCar <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_NorthCarolina/NRI_Table_CensusTracts_NorthCarolina.csv',sep=""), stringsAsFactors = TRUE)
+NRIFlor <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_Florida/NRI_Table_CensusTracts_Florida.csv',sep=""), stringsAsFactors = TRUE)
+EJScreen20 <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/EJSCREEN_2020_StatePctile.csv',sep=""),stringsAsFactors = TRUE)
 #colnames(EJScreen20)

### 2022-04-18 Draft Report/Draft Project.Rmd
Added EJSCreen code and Filtered CSV's (see line 162 and then 200/225/250)

@@
 
+## COMMENT THESE OUT THEN READ THEM IN
+write.table(EJ_California,'./Potential Data Sources/Census Tracts/EJScreen_California.csv',col.names=TRUE,sep=",")
+write.table(EJ_NorCar,'./Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv',col.names=TRUE,sep=",")
+write.table(EJ_Florida,'./Potential Data Sources/Census Tracts/EJScreen_Florida.csv',col.names=TRUE,sep=",")
+###
+
 # EJ_NorCar <- EJScreen20 %>%
@@
 # See GEOID Structure, <https://www.census.gov/programs-surveys/geography/guidance/geo-identifiers.html> all the other stuff is Census Tracts not BlockGroups
+# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_California.csv'
+# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_California.csv'
+# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_California.csv'
 names(EJ_California)[names(EJ_California) == 'ID'] <- 'BLOCKID'
@@
 
+# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_Florida.csv'
+# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_Florida.csv'
+# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_Florida.csv'
 EJScreen_Florida <- EJ_Florida
@@
 
+# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv'
+# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv'
+# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv'
 EJScreen_norcar <- EJ_NorCar
@@
 
+mapview(list(NC_NRI_fire_shp,NC_EJ_shp),
+        zcol=c("WFIR_RISKS","PRE1960PCT"),
+        layer.name=list("Fire Risk","Relative Building Ages"))
+
 # mapshot(m, file = paste0(getwd(), "/map.png"),
